A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}

### 2022-04-04 ICLP Paper/samplepaper.tex
updating based-on reviews

\usepackage{comment}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

In order to counteract limitations of window-wise successive optimization
due to "decomposition mistakes" and unnecessary idle times that do not directly
affect the makespan to some extent,
we have devised two additional techniques that can be incorporated into the
iterative scheduling process.
The first extension is time window \emph{overlapping}, where a configurable percentage
of the operations per time window can still be rescheduled when proceeding
to the next time window.
To this end, an (optimized) partial schedule is postprocessed and the configured
number of operations to overlap are picked in decreasing order of starting times.
For example, if one operation from the first time window is supposed overlap for the
decomposed schedule in Figure~\ref{fig:decomposed}, the second operation of job~\lstinline{2}
whose starting time~\lstinline{4} is latest may be chosen,
which in turn enables its processing after the second operation of job~\lstinline{3}
by rescheduling together with operations of the second time window.
The addition of overlapping operations to the encoding in Listing~\ref{prg:encoding}
requires handling them similar to operations of the time window to schedule,
i.e., enabling the choice of an execution order by the rules in lines
\ref{prg:base:order}-\ref{prg:base:order:end} rather than fixing the order
by the rule in line~\ref{prg:base:ordered}.
As the second extension, we can make use of a stratified ASP program to
postprocess an (optimized) partial schedule by inspecting operations of the
latest time window in the order of starting times whether idle slots
on their machines allow for an earlier execution.
The starting times obtained by the \emph{compression} are then taken instead
of those calculated in the actual schedule optimization, and, e.g.,
the starting time~\lstinline{12} of the third operation of job~\lstinline{1}
would be turned into~\lstinline{9} to fill the idle slot available on
machine~\lstinline{3} when compressing the decomposed schedule displayed in Figure~\ref{fig:decomposed}.

@@

    30\%             & \boldmath{$2943.1$}  & $1000.0$   & $4.0$	   & $3162.2$      & $1000.0$ & $4.0$ & $5857.0$ & $1000.0$ & $9.0$%\\ 
                     % [1.5mm]
    \botline %\hline
    \end{tabular}
   % }
%    \end{minipage}
% }
%    \hfill\hfill\mbox{~}
\end{table}
%
Focusing again on the static M-EST decomposition strategy,
Table~\ref{tab:Table03} assesses different percentages of overlapping
operations from previous time windows as well as the impact of compressing
(optimized) partial schedules in a postprocessing phase.
The baseline approach as above is included in the first row with 0\% of the
operations per time window overlap and without the use of compression.
The next three rows exclude compression as well but introduce 10\%, 20\%, or
30\% of overlapping operations that can be rescheduled together with operations
of the next time window.
In view of the considered number of time windows, picked according to the shortest
average makespan achieved with 0\% overlap, between $18$ and $35$ operations constitute
an overlap of 10\%, so that 30\% range up to more than $100$ operations.
Such non-negligible amounts can make the optimization of subproblems for time windows
harder, while the moderate overlaps of 10\% or 20\% manage to consistently
improve the solution quality.
Even more significant advantages are obtained by applying compression in the last
four rows,
where the average makespan gets already shorter with 0\% overlap.
Further improvements of roughly 2\% in solution quality are achieved with 20\% overlap,
which yield consistently better schedules than the smaller overlap of 10\% and
avoid deteriorations as observed with 30\% overlap on the largest instances
with $100\times20$ jobs and machines.
In summary, a Machine-based problem 
decomposition strategy along with
postprocessing to compress partial schedules as well as a modest overlap of about
20\% of the time window size turn out to be most successful for the JSP
benchmark sets under consideration.

% Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment. Every three sub-columns have illustrated the results of a set of the benchmark instances and each row refers to a decomposition strategy with static (S) or dynamic (D) assignment. The results of this table conclude that the most effective decomposition strategies are machine-based regardless of whether the operations are ordered based on EST or MTWR; however, the dynamic assignment has no positive effect than the static. %The best average makespan of $50\times15$ and $100\times20$ are  $3060.3$, and $5948.6$.
% We applied the overlap between the windows to repair the mistakes generated by the decomposition; in addition, we compressed each optimized window to ensure that all the operations had been started as early as possible. Table \ref{tab:Table03} summarizes the results with/without the compression and the overlap. The first column refers to the overlap percentage where $0\%$ means no overlap between TWs. Every three columns have presented the results of a particular set of instances. We applied the overlap and compression with the solutions of the number of windows, which provided a better average makespan using a machine-based EST strategy. In the non-compression part, the overlap has improved the results slightly through the different sizes where $10\%$ and $20\%$ of overlap provided the best results. However, the compression has shown a significant improvement wrt the running time and the average makespan, especially for the $100\times20$ set. The compression demonstrated effectiveness as it minimized the average makespan and decreased the running time compared to the model without compression. Therefore, the overlap and compression have improved the makespan results by $4.5\%, 3.4\%$ and $6.8\%$ and the running time by $40\%$.

\section{Conclusions}\label{sec:conclusions}
%
Our work develops multi-shot ASP modulo DL methods for JSP solving,
based on problem decomposition into balanced time windows that respect the operation
precedence within jobs and successive makespan optimization while extending
partial schedules to a global solution.
We demonstrate that splitting highly complex JSP instances into
balanced portions, for which partial schedules of good quality can be reliably
found within tight runtime limits, leads to significantly better global solutions
than obtainable with single-shot ASP modulo DL optimization in similar runtime.
Related works taking advantage of single- \citep{abjoossctowa21a}
and multi-shot solving \citep{francescutto2021solving} by
\clingodl to tackle real-world scheduling problems evaluate threshold values
on DL variables and perform optimization by means of common 
\lstinline{#minimize} statements of \clingo rather than minimizing the
value of a DL variable like \lstinline{makespan} in Listing~\ref{prg:encoding}.
Applying a similar approach can be advantageous in our future work as it may
potentially improve computational efficiency as well as solution quality further
and is even a necessity to switch from the makespan as optimization objective to tardiness
w.r.t.\ deadlines, which are crucial in many application areas.
While Machine-based decomposition turned out to be helpful
for finding better schedules, the strategies presented in Section~\ref{subsec:decomposition}
can be regarded as naive heuristics based on straightforward instance properties,
and incorporating machine-learning approaches \citep{elscge22a,tagesc21a}
for more adept partitioning is another promising direction for future work.
Moreover, we aim to generalize our methods further, e.g.,
by considering flexible machine allocations and varying processing times,
and scale them up to large real-world scheduling problems with an order of magnitude more
operations than in synthetic benchmarks.
%
% In this work, different decomposition strategies have been proposed for solving the Job-shop Scheduling Problem (JSP) using a multi-shot Answer Set Programming with Difference Logic constraints to handle the decomposition approach and the problem constraints. The proposed model split the problem into sub-problems using different decomposition strategies and then optimized each separately while minimizing the Makespan. The proposed model has been developed to improve the performance by applying the overlap between TWs to fix the decomposition mistakes. In addition, the compression of each window is to fulfill the gaps that occurred during the optimization process. We tested our model on a set of benchmark instances and the experiments have shown that i) considering the bottleneck machines during the decomposition process is effective, ii) The overlap between TWs improves the solution quality, and iii) the compression of TWs is super practical, as it aided the proposed model to obtain better solutions in less computational time. Our work can be extended by linking the optimization model with a simulator that considers more realistic factors such as machine breakdowns, new job arrivals and order cancellations.

>>>>

\usepackage{comment}
\usepackage{algorithm} 
\usepackage{algpseudocode}
\usepackage{color,soul}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\usepackage[most]{tcolorbox}
\usepackage{caption}

@@

In order to counteract limitations of window-wise successive optimization
due to "decomposition mistakes" and unnecessary idle times that do not directly
affect the makespan to some extent,
we have devised two additional techniques that can be incorporated into the
iterative scheduling process.
The first extension is time window \emph{overlapping}, where a configurable percentage
of the operations per time window can still be rescheduled when proceeding
to the next time window.
To this end, an (optimized) partial schedule is postprocessed and the configured
number of operations to overlap are picked in decreasing order of starting times.
For example, if one operation from the first time window is supposed overlap for the
decomposed schedule in Figure~\ref{fig:decomposed}, the second operation of job~\lstinline{2}
whose starting time~\lstinline{4} is latest may be chosen,
which in turn enables its processing after the second operation of job~\lstinline{3}
by rescheduling together with operations of the second time window.
The addition of overlapping operations to the encoding in Listing~\ref{prg:encoding}
requires handling them similar to operations of the time window to schedule,
i.e., enabling the choice of an execution order by the rules in lines
\ref{prg:base:order}-\ref{prg:base:order:end} rather than fixing the order
by the rule in line~\ref{prg:base:ordered}.
As the second extension, we can make use of a stratified ASP program to
postprocess an (optimized) partial schedule by inspecting operations of the
latest time window in the order of starting times whether idle slots
on their machines allow for an earlier execution.
The starting times obtained by the \emph{compression} are then taken instead
of those calculated in the actual schedule optimization, and, e.g.,
the starting time~\lstinline{12} of the third operation of job~\lstinline{1}
would be turned into~\lstinline{9} to fill the idle slot available on
machine~\lstinline{3} when compressing the decomposed schedule displayed in Figure~\ref{fig:decomposed}. 
\hl{Figure~{\ref{fig:enum}} summarizes the main steps of the Multi-shot solving controlled by a python script.}

\begin{tcolorbox}[title={The main steps of the Multi-shot solving},breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
  \vspace{0.1\baselineskip}
  \begin{enumerate}
    \item[\textbf{1-}] Ground the base program.
    \item[\textbf{2-}] Ground the next shot.
    \item[\textbf{3-}] Optimize the current window till reaching the optima or the timeout.
    \item[\textbf{4-}] Apply the compression of a window (postprocessing phase).
    \item[\textbf{5-}] Keep the operations assignment to be passed to the next window.
    \item[\textbf{6-}] Go to \textbf{2} until all windows are optimized.
  \end{enumerate}
\end{tcolorbox}
\captionof{figure}{Main Steps of \textit{Multi-shot}\label{fig:enum}}

@@

    30\%             & \boldmath{$2943.1$}  & $1000.0$   & $4.0$	   & $3162.2$      & $1000.0$ & $4.0$ & $5857.0$ & $1000.0$ & $9.0$%\\ 
                     % [1.5mm]
    \botline %\hline
    \end{tabular}
   % }
%    \end{minipage}
% }
%    \hfill\hfill\mbox{~}
\end{table}
%
Focusing again on the static M-EST decomposition strategy,
Table~\ref{tab:Table03} assesses different percentages of overlapping
operations from previous time windows as well as the impact of compressing
(optimized) partial schedules in a postprocessing phase.
The baseline approach as above is included in the first row with 0\% of the
operations per time window overlap and without the use of compression.
The next three rows exclude compression as well but introduce 10\%, 20\%, or
30\% of overlapping operations that can be rescheduled together with operations
of the next time window.
In view of the considered number of time windows, picked according to the shortest
average makespan achieved with 0\% overlap, between $18$ and $35$ operations constitute
an overlap of 10\%, so that 30\% range up to more than $100$ operations.
Such non-negligible amounts can make the optimization of subproblems for time windows
harder, while the moderate overlaps of 10\% or 20\% manage to consistently
improve the solution quality.
Even more significant advantages are obtained by applying compression in the last
four rows,
where the average makespan gets already shorter with 0\% overlap.
Further improvements of roughly 2\% in solution quality are achieved with 20\% overlap,
which yield consistently better schedules than the smaller overlap of 10\% and
avoid deteriorations as observed with 30\% overlap on the largest instances
with $100\times20$ jobs and machines.
In summary, a Machine-based problem 
decomposition strategy along with
postprocessing to compress partial schedules as well as a modest overlap of about
20\% of the time window size turn out to be most successful for the JSP
benchmark sets under consideration.\\

% Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment. Every three sub-columns have illustrated the results of a set of the benchmark instances and each row refers to a decomposition strategy with static (S) or dynamic (D) assignment. The results of this table conclude that the most effective decomposition strategies are machine-based regardless of whether the operations are ordered based on EST or MTWR; however, the dynamic assignment has no positive effect than the static. %The best average makespan of $50\times15$ and $100\times20$ are  $3060.3$, and $5948.6$.
% We applied the overlap between the windows to repair the mistakes generated by the decomposition; in addition, we compressed each optimized window to ensure that all the operations had been started as early as possible. Table \ref{tab:Table03} summarizes the results with/without the compression and the overlap. The first column refers to the overlap percentage where $0\%$ means no overlap between TWs. Every three columns have presented the results of a particular set of instances. We applied the overlap and compression with the solutions of the number of windows, which provided a better average makespan using a machine-based EST strategy. In the non-compression part, the overlap has improved the results slightly through the different sizes where $10\%$ and $20\%$ of overlap provided the best results. However, the compression has shown a significant improvement wrt the running time and the average makespan, especially for the $100\times20$ set. The compression demonstrated effectiveness as it minimized the average makespan and decreased the running time compared to the model without compression. Therefore, the overlap and compression have improved the makespan results by $4.5\%, 3.4\%$ and $6.8\%$ and the running time by $40\%$.
\hl{We conducted other experiments to assess the quality of the solutions obtained by our proposed method by comparing ASP with single-shot and OR tools. Table~{\ref{tab:Table04}} summarizes the results of our experiments. The values in the table represent the average distance to the optimal solutions for each set of instances. The first column represents the size of the instances with the decomposition parameters. The second and third columns show the average distance of the results obtained by ASP single and multi-shot, respectively. The last column shows the average distance obtained by OR tools. The results show that the single-shot ASP is far from the optima, especially with the largest dataset. In addition, it is shown that the developed model improves the results compared to the pure ASP. However, the performance of OR tools is better than ours, especially for the first two sets of instances. For $100 \times 20$, our proposed model is a bit comparable to OR tools, which means that if the instance size increases, it could be possible to obtain better results with our approach.}

\begin{table}[t]
  \caption{Experiments comparing the Single, Multi-shot ASP and OR tools\label{tab:Table04}}%
%    \mbox{~}\hfill
%    \resizebox{0.8\textwidth}{!}{%
%    \begin{minipage}{\textwidth}
  \centering
  \begin{tabular}{l l r r }

  \topline%\hline
  %Timeout & \multicolumn{2}{c}{$1000$ sec} & \multicolumn{2}{c}{$500$ sec} \\
   & \multicolumn{2}{c}{ASP} & OR tools \\
  Instance set & Single-shot & Multi-shot &
  \midline

  $50\times15$  & $768.3$ & $169.3$ & $5.0$ \\

  %$50\times15$  & $169.3$ & $5.0$ & $163.3$ & $26.0$ \\
  ($4$w, M-EST, 30\%) & & & \\

  $50\times20$  & $1035.5$ & $303.8$ & $70.4$ \\
  %$50\times20$  & $303.8$ & $70.4$ & $304.4$ & $86.2$ \\
  ($4$w, M-EST, 20\%) & & & \\

  $100\times20$  & $41420.4$ & $253.5$ & $193.6$ \\
  %$100\times20$  & $253.5$ & $193.6$ & $286.3$ & $284.1$ \\
  ($9$w, M-EST, 20\%) & & &

  \botline %\hline
  \end{tabular}
\end{table}
%

\section{Conclusions}\label{sec:conclusions}
%
Our work develops multi-shot ASP modulo DL methods for JSP solving,
based on problem decomposition into balanced time windows that respect the operation
precedence within jobs and successive makespan optimization while extending
partial schedules to a global solution.
We demonstrate that splitting highly complex JSP instances into
balanced portions, for which partial schedules of good quality can be reliably
found within tight runtime limits, leads to significantly better global solutions
than obtainable with single-shot ASP modulo DL optimization in similar runtime.
Related works taking advantage of single- \citep{abjoossctowa21a}
and multi-shot solving \citep{francescutto2021solving} by
\clingodl to tackle real-world scheduling problems evaluate threshold values
on DL variables and perform optimization by means of common 
\lstinline{#minimize} statements of \clingo rather than minimizing the
value of a DL variable like \lstinline{makespan} in Listing~\ref{prg:encoding}.
Applying a similar approach can be advantageous in our future work as it may
potentially improve computational efficiency as well as solution quality further
and is even a necessity to switch from the makespan as optimization objective to tardiness
w.r.t.\ deadlines, which are crucial in many application areas.
Moreover, we aim to generalize our methods further, e.g.,
by considering flexible machine allocations and varying processing times,
and scale them up to large-scale instances with an order of magnitude more
operations than in synthetic benchmarks.
%
% In this work, different decomposition strategies have been proposed for solving the Job-shop Scheduling Problem (JSP) using a multi-shot Answer Set Programming with Difference Logic constraints to handle the decomposition approach and the problem constraints. The proposed model split the problem into sub-problems using different decomposition strategies and then optimized each separately while minimizing the Makespan. The proposed model has been developed to improve the performance by applying the overlap between TWs to fix the decomposition mistakes. In addition, the compression of each window is to fulfill the gaps that occurred during the optimization process. We tested our model on a set of benchmark instances and the experiments have shown that i) considering the bottleneck machines during the decomposition process is effective, ii) The overlap between TWs improves the solution quality, and iii) the compression of TWs is super practical, as it aided the proposed model to obtain better solutions in less computational time. Our work can be extended by linking the optimization model with a simulator that considers more realistic factors such as machine breakdowns, new job arrivals and order cancellations.
